practical machine learning project write up
===========================================
##### author : godspeed


### 1.packages
```{r,results="hide"}
library(caret)
library(dplyr)
library(randomForest)
```

### 2.getting data
##### information is available from the website : 
##### http://groupware.les.inf.puc-rio.br/har
##### training data :
##### https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
##### test data :
##### https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
setwd("~/Desktop")
# data contains "DIV/O!"" and "NA", read them as NA
data_train<-read.csv("pml-training.csv",header=T,na.strings=c("NA","#DIV/0!",""))
data_test<-read.csv("pml-testing.csv",header=T,na.string=c("NA","#DIV/0!",""))
```

### 3.objective
##### predict the manner in which they did the exercise. This is the "classe" variable in the training set. Use any of the other variables to predict with.
```{r}
table(data_train$classe)
```

### 4.cleaning data
##### remove the variables : 
##### 1.does not make sense(column 1 to 5)
##### 2.near zero vriance
##### 3.columns that contains a lot NAs(90%)
```{r}
# remove first five column
data_train <- data_train[, -(1:5)]
# remove nearly zero
index_nsv <- nearZeroVar(data_train)
data_train <- data_train[, -index_nsv]
# remove columns contains a lot NA, keep low NA frequency column
index_NA_90percent <- sapply(data_train, function(x) mean(is.na(x))) > 0.9
data_train <- data_train[, index_NA_90percent==F]
```


### 5.split train data into training set (80%) and testing set (20%). 

```{r}
set.seed(1)
inTrain <- createDataPartition(data_train$classe, p = 0.8, list = FALSE)
set_training <- data_train[inTrain,]
set_testing <- data_train[-inTrain,]
```


### 6.model training with 3 cross validation
```{r}
train_control <- trainControl(method="cv", number=3)
model <- train(classe~.,data=set_training, method="rf",trControl=train_control)
```

### 7.prediction on training set and testing set
```{r}
# error rate in sample
prediction <- predict(model, set_training, type="raw")
error_rate_insample<-confusionMatrix(prediction, set_training$classe)

print(paste("error rate in sample : ",1-error_rate_insample[["overall"]][["Accuracy"]]))

error_rate_insample

# error rate out of sample
prediction <- predict(model, set_testing, type="raw")
error_rate_outsample<-confusionMatrix(prediction, set_testing$classe)

print(paste("error rate out of sample : ",1-error_rate_outsample[["overall"]][["Accuracy"]]))

error_rate_outsample
```

### 8.prediction on test data
```{r}
prediction_test <- predict(model, data_test, type = "raw")
prediction_test
```

### 9.submission
```{r}
answers <- prediction_test

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)

```